Playing with PCAs

### 2024-07-29 3-reports/playing_with_pcas.Rmd
Update playing_with_pcas.Rmd

pca_owc = fit_pca_function(data_wide_PCA %>% filter(site == "OWC")) 
pca_ptr = fit_pca_function(data_wide_PCA %>% filter(site == "PTR")) 
pca_crc = fit_pca_function(data_wide_PCA %>% filter(site == "CRC")) 
pca_gcw = fit_pca_function(data_wide_PCA %>% filter(site == "GCW")) 
pca_msm = fit_pca_function(data_wide_PCA %>% filter(site == "MSM")) 
pca_gwi = fit_pca_function(data_wide_PCA %>% filter(site == "GWI"))

@@

@@

df = data_wide_PCA %>% filter(region == "CB") %>%   drop_na() %>% dplyr::select(where(is.numeric))
grp = data_wide_PCA %>% filter(region == "CB")%>% drop_na() %>% dplyr::select(!where(is.numeric))

@@

@@

>>>>

pca_owc = fit_pca_function(data_wide_PCA2 %>% filter(site == "OWC")) 
pca_ptr = fit_pca_function(data_wide_PCA2 %>% filter(site == "PTR")) 
pca_crc = fit_pca_function(data_wide_PCA2 %>% filter(site == "CRC")) 
pca_gcw = fit_pca_function(data_wide_PCA2 %>% filter(site == "GCW")) 
pca_msm = fit_pca_function(data_wide_PCA2 %>% filter(site == "MSM")) 
pca_gwi = fit_pca_function(data_wide_PCA2 %>% filter(site == "GWI"))

@@

pc_region_transect %>% 
  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
  ggplot(aes(x = abs(PC1), y = variable, color = transect))+
  geom_point(size = 5)+
#  scale_colour_manual(values = c("red", "blue"))+
 # xlim(0, 0.4)+
  #geom_vline(xintercept = 0)+
  labs(y = "",
       x = "|PC1|")+
  facet_grid(. ~ region)+
  theme_kp()+
  geom_vline(xintercept = 0.3, linetype = "dashed")+
  geom_vline(xintercept = 0.1, linetype = "dashed")

@@

df = data_wide_PCA2 %>% filter(region == "WLE") %>%   drop_na() %>% dplyr::select(where(is.numeric))
grp = data_wide_PCA2 %>% filter(region == "WLE")%>% drop_na() %>% dplyr::select(!where(is.numeric))

@@

ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1,
                   groups = as.character(pca_cluster$grp$cluster), 
                   ellipse = F, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_convexhull(aes(group = pca_cluster$grp$cluster,
                      fill = pca_cluster$grp$cluster), alpha = 0.2)+
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_cluster$grp$transect,
                 color = groups))+ 
  # scale_color_manual(values = c("#16879C", "#BB281E" ))+
  #scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  # xlim(-4,20)+
  # ylim(-8,8)+
  labs(fill="cluster",
       color = "cluster",
       shape = "region",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

@@

## PCA on C-normalized data

```{r}
data_wide_PCA2 = 
  data_wide_PCA %>% 
  pivot_longer(cols = c(where(is.numeric), -TC)) %>% 
  mutate(value = value/TC) %>% 
  dplyr::select(-TC) %>% 
  pivot_wider()

pca_all_c = fit_pca_function(data_wide_PCA2)

ggbiplot(pca_all_c$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_all_c$grp$region), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(#shape = pca_all_c$grp$site,
                 color = groups))+ 
  scale_color_manual(#breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  #scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  xlim(-6,6)+
  ylim(-4.5,4.5)+
  labs(shape="",
       title = "PCA: all, normalized to C",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

```{r}
df = data_wide_PCA2 %>%   drop_na() %>% dplyr::select(where(is.numeric))
grp = data_wide_PCA2 %>% drop_na() %>% dplyr::select(!where(is.numeric))

library(ggConvexHull)

df_cl %>% 
  ggplot(aes(x = transect, y = cluster, color = site))+
  geom_jitter(width = 0.1, height = 0.1, size = 3)+
  facet_wrap(~region)+
  theme_kp()
```
